\appendix
\newgeometry{margin=2cm}
% change chapter title to be without word chapter
% http://www.latex-community.org/forum/viewtopic.php?f=4&t=638
% appendicies do not need space in beginning of the chapter
\makeatletter
\renewcommand{\@makechapterhead}[1]{%
	\vspace*{1 pt}%
	{\color{chapters}\setlength{\parindent}{0pt} \raggedright \normalfont
	\bfseries\Huge\thechapter\ #1
	\par\nobreak\vspace{1 pt}}}
\makeatother

\input{appendix-cert-letter}

\input{./lab/preliminary-test}

\input{preliminary-course-gnu-linux}

\input{appendix-1}

\input{appendix-2}

\input{appendix-3}

\chapter{Feedback from students}
\section{Comments from intensive course students}



\subsubsection{Feedback from second group: International students}
Intensive Programme 2013 "Deploying IT Infrastructure Solutions" contains one hands-on course: Protecting web applications. Students were from Estonia, Finland, Greece and from Lithuania. 
Only Estonian students (1/3 from all students) had previous experience with GNU/Linux and this practical class was too hard for most of the students because insufficient previous knowledge:

 Comments from students (unchanged): 
*It was a bit fast, but as I have been dealing with DVWA and this kind of course before, it was okay for me. I was helping other participants. 
*quite hard to follow and was specific for one team. 
*Very important for the Security team, and good new perspectives and experiences for everyone else too. 
*Interesting course. Liked that Estonians helped people who weren’t very familiar with linux. Nothing new to me. 
*This was carried out as a practical class and was quite tough since I had little previous Linux experience. 
*It was a very long, tiring practice lecture. Somehow it would have been better to separate the long lecture into several parts. It would have been more productive. Otherwise tired lecturer and students couldn’t follow.  
*I took my first baby steps towards being professional hacker. In some cases, following teachers teaching was hard because of the fast pace. 
*Not that relevant for all projects, but still important thing to understand as a IT student. 
*Awesome, hacking is interesting. 
*Nice topic and useful to our project. It was like a first bite of our project. And Margus taught things so that everyone can learn.
*That was something different out of software engineering field 
*Very good lecture. Difficult to follow. 
*I really like to search for security holes and vulnerability threats in web pages and this tool is one of the best for testing, but I didn't manage to follow the teacher's step by step tutorial. I have got messed up in virtual machines, because I didn't have the experience for them, however I have managed to learn more about them after this topic, also Margus explained more about DVWA tool for our team, so I got the idea.

\chapter{Lab proposals for the future}
\label{Appendix:Lab proposals for the future}

 
 \begin{enumerate}
   \item E-mail services (4 days)
    \begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
  		\item SPAM control
	  	\item Virus protection
  		\item MTA's 
	  	\item MDA's
    \end{enumerate}
    \item IP firewalls and IDS/IPS (4 days)
        \begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
  		\item IP firewalls netfilter/iptables and packet filter (pf) (2 days)
	  	\item IDS/IPS (2 days)
  		\item NetFlow (together with CERT.EE)
		\end{enumerate}
    \item Autentication and authorization (4 days)
        \begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
  		\item LDAP and Samba4 AD.To pass student needs to configure authentication service using SAMBA4 with Zentyal server and join GNU/Linux workstation into domain using LikeWise Open plugin.
	  	\item Windows and Linux clients with Samba4 AD 
  		\item Web application authentication with Samba4 AD and LDAP
    		\end{enumerate}
    \item GNU/Linux central management with Puppet
        \begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
	  		\item Installation of Puppet using passenger (2 day)
		  	\item Writing puppet recipes (1 day)
    		\end{enumerate}
    	\item Central logging (3 days)
    	    \begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
	  		\item Collecting logs with rsyslog/syslog-ng (1 day)
		  	\item Monitor and analyse log files (1 day)
    		\end{enumerate}
    		\item Cloud and virtualization solutions
    		\begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
	  		\item ProxMox, libvirtd, \gls{KVM}, virtual networks
		  	\item Private cloud for fileserver OwnCloud
		  	\item Open source cloud platform (OpenStack or OpenCloud or Eucalyptus)
    		\end{enumerate}
 \end{enumerate}





\chapter{Toormaterjal}
{\color{red} See ei ole osa tööst vaid asjad, mis jäävad tõenäoliselt välja}


Siin on asjad, kust tõstan mingid lõigud töösse

\begin{verbatim}
8.8.8.8; pwd

8.8.8.8; cp ../../config/config.inc.php /tmp/asi.txt

8.8.8.8; cat /tmp/asi.txt

Kuigi ping tulemuste all pole config faili sisu näha, 
tehes lehel parema kliki ja valides "View Page Source" näeme, et source faili sees on config.inc.php sisu olemas.


    DVWAs kasutades "Command execution" andmebaasi loomine:
    8.8.8.8; mysql --user=root --password='student' -e 'create database kala;'

Software eggs
http://www.itcollege.ee/?=PHPB8B5F2A0-3C92-11d3-A3A9-4C7B08C10000
    

\end{verbatim}

\url{http://blog.spiderlabs.com/2011/01/detecting-malice-with-modsecurity-csrf-attacks.html}

Lisada upstart ülesanne mysql oom score tuunimiseks.


DHCP laborisse lisada shared-network konfig

Iga peatüki juurde panna kirja, mitu lehekülge on vaja.
E-kiri tsitaatidena - ainult vajalikud osad.
ainekava ei tõlgi - Jääb lisasse eesti keelsena
neljapäeva hommikul 16. saadan töö. - tühja kohta ei ole - todo on tehtud ja jääb keelekorrektuur.
neljapäevaks kaitsmisproov.

 Todays studi in \gls{EITC} include lectures, practical classes and independent work as homework. By and large, one subject is divided as follows: 25\% lectures, 25\% practical classes and 50\% homework which is mostly working with materials (books, web articles). In the worst case, practical work constitutes only 25\% and takes place at \gls{EITC} computer classes.
Moreover, the students are not interested in learning mere theory. The formulas are not seen necessary nor linked to their study area or future job. Theory that is not used will be forgotten quickly. In a few years students won't even remember if a specific topic was covered or not. Applied education should introduce practical approach and learning by doing. In ITC field practical classes and practical homework is the key to achieving acceptable results.
In initial investigation phase the authors role was curricula development, requirements management,  planning hands-on labs and course descriptions.

{\color{red} 
Usable to design a learner centric course instead of teacher centric. 

\url{http://www.youtube.com/watch?v=zB92UMyYzKM}

Course content - get attencion
Ingeraction amoungs students
Jugement 
Student diversity - ebaühtlus
Create a role played scenario
Analyse the material.

\url{http://www.youtube.com/watch?v=oRgLqEF-qAU}

}

Serdi sees url -  Online Certificate Status Protocol (OCSP)
Saab teha ajaaknaga harjutusi. Teenuste uuendamine ja uure vahendite kasutamine ründamiseks...Kõik oleks ajas muutuv.

\url{http://www.createdebate.com/debate/show/How_relevant_is_the_ADDIE_model_in_2009}

ADDIE model \url{http://www.youtube.com/watch?v=fdpHO1xycgo&list=PLEBAA915299F438EB}


Haridustehnoloogia sõnastik - \url{http://wiki.e-uni.ee/htsonastik/index.php?n=Main.Otilde#oppematerjal}

Miks creative common - \url{http://www.copyrightreform.eu/case-for-copyright-reform}

Alternative Training Models
Advances in Developing Human Resources November 2006 8: 460-475,
Today a learning by doing approach is accepted way to gain new skills and knowledge in cyber security field. This thesis focuses to develop of the practical hands-on e-course for system administrators in \gls{EITC}. Moreover the developed laboratories are used in higher education and also in continuous education classes.

%\input{test}