\chapter{Future Research}
\label{Future Research}

\lettrine[lraise=0.1, nindent=0em, slope=-.5em]{\color{Violet}F}{uture} research...


\begin{itemize}
\item Ainult süsteemide halduse oskusest ei piisa, peab oskama kommunikeeruda, intsidente uurida jne...
\item IPv6 implementeerimine erinevates juhendites ja laborites
\end{itemize}
{\color{red} Väheke puudujääke ja probleeme, siis ideed ja mõtte, mida tahaks veel teha
Not all developed learning materials are following consistent style and using \gls{git}
Blue Teams were challenged by additional Developers


tasks and media pressure, requiring them to have a wide range of skills to be successful. 


Locked Shields should remain as a live technical cyber defence exercise, as there is a clear need for
more and similar training events. In this report, we have listed numerous observations as to how to
improve Locked Shields in the future.

Jätame välja NFS, NIS unsecure LDAP jne ründed.
Samuti jätame välja password crackingu. pigen nõuame, et rakendataks OS PAM password policyt + cracklibi.
 }
 
 
File services  - private public cloud näiteks OwnCloud should be covered.
 
 \begin{enumerate}
 
     \item E-mail services (4 days)
    \begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
  		\item SPAM control
	  	\item Virus protection
  		\item MTA's 
	  	\item MDA's
    \end{enumerate}
    \item IP firewalls and IDS/IPS (4 days)
        \begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
  		\item IP firewalls netfilter/iptables and packet filter (pf) (2 days)
	  	\item IDS/IPS (2 days)
  		\item NetFlow (kuna seda loetakse CERT.EE abil, siis jääb välja)
		\end{enumerate}
    \item Autentication and authorization (4 days)
        \begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
  		\item LDAP and Samba4 AD.To pass student needs to configure authentication service using SAMBA4 with Zentyal server and join GNU/Linux workstation into domain using LikeWise Open plugin.
	  	\item Windows and Linux clients with Samba4 AD 
  		\item Web application authentication with Samba4 AD and LDAP
    		\end{enumerate}
    \item GNU/Linux central management with Puppet
        \begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
	  		\item Installation of Puppet using passenger (2 day)
		  	\item Writing puppet recipes (1 day)
    		\end{enumerate}
    	\item Central logging (3 days)
    	    \begin{enumerate}[label=LAB \arabic*.,leftmargin=*]
	  		\item Collecting logs with rsyslog/syslog-ng (1 day)
		  	\item Monitor and analyse log files (1 day)
    		\end{enumerate}
 \end{enumerate}
 